Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms

### 2019-07-04 manuscript/main.tex
Add updated manuscript text

\section*{Tasks} 
\begin{enumerate}
\item What is novel about this? The first anyone has treated an ecosystem like an
organism. The first time anyone has done this correlative approach well
by considering environmental and spatial structure in the data. 
\item Figure out variance partitioning (with Hannah)
\item Maybe variance partitioning plus correlations (varPart)
\item Do PC correction in all combinations
\item Identify significant taxa and build a venn diagram figure (or
something)
\item 
\end{enumerate}
\section{Introduction}

I want to know whether microbial community composition is important for determining ecosystem-scale functions independent of the underlying environmental variation.

Microorganisms mediate many biogeochemical processes, but despite 
much effort to link microbial communities
to these functions the mechanism by which they regulate these
processes remains unknown. Ecologists have investigated the effects of changing
biodiversity on ecosystem function for decades, but many of their approaches
are not applicable in microbial ecology due to the nature of microbial data. 
This limits our ability to predict and manage globally relevant biogeochemical cycles. To
address this challenge, we propose integrating current theory on community
assembly and the functioning of ecosystems with lessons from genotype-phenotype
mapping in quantitative genetics. This will address the
question of how does variation in microbial community structure contribute to
variation in the rate of ecosystem function?

>>>>

\section*{Tasks} 
\begin{enumerate}
\item What is novel about this? The first anyone has treated an ecosystem like an
organism. The first time anyone has done this correlative approach well
by considering environmental and spatial structure in the data. 
\item Interpret model output
\item Decide whether to include Gabon data, Brazil data, or both
\end{enumerate}

\section{Introduction}

Microorganisms mediate many biogeochemical processes and yet the link between
microbial biodiversity and ecosystem function remains unknown. 
Ecologists have investigated the effects of changing
biodiversity on ecosystem function for decades, but this research has primarily
focused on linking terrestrial plant community richness with productivity
\citep{hooper2005}. More recently, there has been an interest in linking
microbial community composition to ecosystem functions such as host disease 
and soil greenhouse gas emissions \citep{schimel1998, singh2010}. 
However, it remains unclear to what extent 
microbial community composition is important for determining the rates of ecosystem-scale
functions.  This uncertainty limits our ability to predict and manage host
health and global biogeochemical cycles. To
address this challenge, we propose integrating traditional
biodiversity-ecosystem function research with ideas from genotype-phenotype
mappingin population genetics. This connection will help determine
whether microbial community
composition alters ecosystem-scale functions independent
of the underlying environmental variation.

If local selection always optimizes the available biodiversity to maximize
ecosystem function, them microbial community composition should not matter for
pedicting the rates of microbially mediated functions. Instead, ecosystem function
should be predicted from the underlying environmental variation. In this case,
the microbial community simply acts as a conduit through which the abiotic
environment alters ecosystem function. However, experiments that manipulate the
connection between environmental factors and microbial community composition
such as environmental treatments, common gardens, and reciprocal transplants
have shown different rates of ecosystem functions for different microbial
communities. This has been observed for plant decomposition
\citep{strickland2009}, plant phenology \citep{panke-buisse2015}, soil nitrogen
cycling \citep{balser2005}, and soil greenhouse gas emissions
\citep{cavigelli2000}. Therefore, variation in microbial community composition
seems to generate variation in ecosystem function. However, in order to
incorporate microorganisms into ecosystem models, we need to identify what
aspects of microbial biodiversity are associated with variation in ecosystem
function.

Most studies of microbial community function focus on one of two aspects of
microbial community structure that are hypothesized to be predictive of
ecosystem function. The first aspect is functional gene or transcript abundance.
In this case, qPCR or shotgun metagonemic sequencing is used to target a gene
or transcript that is a putative marker for a microbial process. For example,
the gene \textit{mcrA}, which encodes the enzyme that performs the final step in
methanogenesis, is hypothesized to be correlated with the rate of
methane production. Other examples include \textit{pmoA} and methanotrophy,
\textit{nifH} and nitrification, and \textit{nosZ} and denitrification. The
rational for this hypothesis is that the abundance of a functional group (as
inferred from marker gene abundance) should correspond to the rate of the
process performed by that group. There are examples where this relationship
holds \citep{freitag2009, freitag2010, schnyder2018}. However, a review of these
studies found that the abundance of a functional gene or transcript is rarely
correlated with the rate of the corresponding process, with most effects either
negative or not significant \citep{rocca2015}.

The second aspect of microbial community structure hypothesized to predict
ecosystem function is taxonomic or functional
diversity. Diversity is either estimated using a measure of community
composition such as 16S rRNA gene ribotyping or manipulated through some proxy
of diversity such as sequential dilution or varying filter sizes. Studies that
experimentally manipulate diversity generally find a relationship between
the applied diversity treatment and ecosystem functions including methanotrophy
\citep{schnyder2018}, phosphorus leaching, greenhouse gas emissions
\citep{wagg2014}, and decomposition \citep{maron2018}. However, in each of these
cases, richness is confounded with other factors such as presence or absence of
major phylogenetic groups (nematodes, fungi) or abundance of microbial cells. In
addition, assembled microbial communities pose the same challenges as
macroorganismal diversity experiments, for example no replication at the highest
diversity treatment. A review of microbial diversity studies shows that
microbial community taxonomic and functional diversity add little variance
explained to models of ecosystem function. Overall, functional gene and
abundance and community doversity improve models of ecosystem function less than
one third of the time and increase variance explained by only 8 percentage
points \citep{graham2016}.

These approaches follow on the tradition of macroorganismal
biodiversity-ecosystem function research by measuring or manipulating diversity
and comparing it to function or by looking at the association between richness
or abundance of a functional group (e.g. plants) and the function performed by
that group (e.g. productivity). However, in the case of microbes, our knowledge
of functional groups is quite limited. We cannot be sure that we're accurately
estimating the abundance or diversity of soil microbial functional groups the
way we can, say, plants in a grassland.

Organisms exhibit population differentiation due to geographic and ecological
processes \citep{wright1943}. Neutral processes such as local genetic drift and
reduced gene flow due to geographic separation can lead to population
differentiation. In addition, environmental distance can lead to selection
against migrants \citep{hendry2004} and local selection on endemic populations
to generate variation. Microorganisms exhibit biogeographic patterns due to
ecological selection and geographic separation \citep{martiny2006}. If dispersal
is non-limiting then the most productive species should always be present and
composition should not matter for local ecosystem function. However, if important
functional groups are dispersal-limited, lost to drift, or are unable to
colonize a site, then there could be important functional consequences of
community differentiation to geographic or ecological isolation.

There are a numbber of ways that environmental or spatial distance could obscure
microbial BEF relationships. For example, historical effects of geographic
isolation such as dispersal limitation could lead to spurious associations
between the abundance of taxa and the rate of ecosystem function. Community
stratification where taxa co-occur due to historical effects or local selection.
Environmental selection can also shape microbial community
structure. For example, pH is known to regulate both broad biodiversity patterns for
microbial communities \citep{fierer2006} and methane cycling dynamics
\citep{ye2012}. This could like to microbial taxa that are correlated with
methane cycling, but are mechanisticalyl unrelated. In fact, if the rate of
ecosystem function is high in one set of ecosystems, function will be
associated with any taxa that are abundant in that set of
ecosystems.

In order to understand the biological basis of ecosystem function, we need to
separate the close association between environmental variation, community
composition, and ecosystem function in order to identify what aspects of
microbial biodiversity determine the rate of ecosystem function.

The challenge of mapping microbial community structure to ecosystem function is
analogous to the challenge of identifying the genetic basis of complex traits in
organismal populations. In the latter case, a population exhibits variation in a
phenotype (e.g. hight or disease state) as well as variation in potentially
thousands of genetic loci (usually identified as single nucleotide polymorphisms
(SNPs)). To identify the genetic basis for a trait, investigators sample from
this population and correlate phenotype with genotype. In most cases, there are
many more loci than individuals and we do not know the function of genes
associated with these SNPs. Genetic markers identified in these genome-wide association
studies can be the target for future research to dissect the function of those
genes, can be added or removed to test their association with the trait, or can
be used as marker for that disease or phenotype even if it has no known
mechanistic association.

This mirrors the problem of microbial
structure-function mapping. In this case, a population of hosts or ecosystems exhibit
variation in a particular trait or ecosystem function putatively mediated by
microorganisms as well as variation in metagenomic composition. Investigators
could sample from this population of ecosystems and correlate the rate of an
ecosystem function with microbial community composition or gene composition.

Population geneticists trying to make robust associations between genetic loci
and phenotype must take into account population stratification. Population
stratification due to shared ancestry or local selection can lead to spurious
associations between genes and traits. For example, human height varies
across Europe in a latitudinal gradient. Positive associations between genes and
phenotypes can arise because of true associations, through linkage
disequilibrium with true genes, or through admixture \citep{lander1994}.
Microbial structure-function connections could pose these same problems. There
could be true associations between a microorganism and the ecosystem function
performed by that taxon, or there could be there could be systematic
co-occurrence patterns between one taxon and a causal taxon like metabolic
networks, or there could be admixture due to shared history or selection within
a sub-set of ecosystems. Robust structure-function associations rely on
controlling for these factors.
